---
title: "R Notebook"
output: html_notebook
---

# 4.4 For-loops

sum of 1+2+... +n is n(n+1)/2
What if we are not sure?

```{r}
compute_s_n <- function(n) {
  x <- 1:n
  sum(x)
}
  
```

how can we compute Sn for various values of n?
We us for-loops here: performing same task over and over, changing value of n

```{r}
for(i in 1:5){
  print(i)
}
```

simple example of for-loop

```{r}
m <- 25
s_n <- vector(length = m) # create an empty vector
for(n in 1:m){
  s_n[n] <- compute_s_n(n)
}
```

in each iteration n=1, ... we compute Sn and store it in the nth entry of s_n

```{r}
n<- 1:m
plot(n, s_n)
```

quadratic; you're right
n(n+1)/2

```{r}
head(data.frame(s_n = s_n, formula = n*(n+1)/2))
```

we can also overlay two results by using function 'lines'

```{r}
plot(n, s_n)
lines(n, n*(n+1)/2)
```

# 4.5 Vectorization and functionals

we rarely use for-loops in R
vectorization is preferred - results in shorter, clearer code

```{r}
x<- 1:10
sqrt(x)
y <- 1:10
x*y
```

not all functions work this way
compute_s_n does not work element-wise

```{r}
n <- 1:25
compute_s_n(n)
```

functionals; functions that help us apply the same function to each entry in a vector, matrix, data frame or list
'sapply' perform element-wise operations on any function

```{r}
x <- 1:10
sapply(x, sqrt)
```

each element of x is passed on to the function sqrt
the result is a vector of the same length as the original x
for-loops can be written as:


```{r}
n <- 1:25
s_n <- sapply(n, compute_s_n)
plot(n, s_n)
```

Other functionals are apply, lapply, tapply, mapply, vapply, and replicate


# 4.6 Exercises

Q1. Return?  
```{r}
x <- c(1,2,-3,4)

if(all(x>0)){
  print("All Postives")
} else{
  print("Not all positives")
}
```
A1. Not all positives

Q2. Which of the following expressions is always FALSE when at least one entry of a logical vector x is TRUE?

A. all(x)
B. any(x)
C. any(!x)
D. all(!x)

A2. D

Q3. The function nchar tells you how many characters long a character vector is. Write a line of code that assigns to the object new_names the state abbreviation when the state name is longer than 8 characters.

A3. 
```{r}
library(dslabs)
data(murders)
new_names <- ifelse(nchar(murders$state)>8, murders$abb, murders$state)
new_names
```

Q4. Create a function sum_n that for any given value, say n, computes the sum of the integers from 1 to n (inclusive). Use the function to determine the sum of integers from 1 to 5,000.

```{r}
sum_n <- function(n) {
  x <- 1:n
  sum(x) 
}

```

```{r}
sum_n(5000)
```

Q5. Create a function altman_plot that takes two arguments, x and y, and plots the difference against the sum.

A5. 
```{r}
altman_plot <- function(x, y) {
plot(x+y, abs(x-y))  
}
```

Q6. After running the code below, what is the value of x?

```{r}
x <- 3
my_func <- function(y){
  x <- 5
  y+5
}
```

A6. 3

Q7. Write a function compute_s_n that for any given n computes the sum 1^2+ .... n^2. report the value of the sm when n=10

```{r}
compute_s_n <- function(n){
  x<- 1:n
  sum(x^2)
}
```

```{r}
compute_s_n(10)
```

Q8. Define an empty numerical vector s_n of size 25 using s_n <- vector("numeric", 25) and store in the results of s1, s2, ...s25 using a for-loop

```{r}
s_n <- vector("numeric", 25)
for(i in 1:25){
  s_n[i] <- compute_s_n(i)
}
s_n
```

Q9. Repeat exercise 8, but this time use sapply.

```{r}
x<- 1:25
sapply(x, compute_s_n)
```

Q10. Repeat exercise 8, but this time use map_dbl.

```{r}
?map_dbl
library(purrr)
map_dbl(n,compute_s_n)
```

Q11. plot Sn versus n

```{r}
plot(n, s_n)
```

Q12. confirm that the formula for this sum is Sn=n(n+1)(2n+1)/6

```{r}
plot(n, s_n)
lines(n, n*(n+1)*(2*n+1)/6)
```
