---
title: "R Notebook"
output: html_notebook
---

# Chapter 4 Programming basics

R is a programming language: can develop complex package. But we introduce three key programming concepts: conditional expressions, for-loops and functions. widely used functions but not cover in this book - split, cut, do.call, Reduce

# 4.1 Conditional Expressions

used for 'flow control'
The most common conditional expression ; if-else statement

general structure of an if-else statement

```{r}
a <- 0
if(a!=0){
  print(1/a)
} else{
    print("No reciprocal for 0.")
}

```

```{r}
library(dslabs)
data(murders)
murder_rate <- murders$total / murders$population*100000
```

which states have a murder rate lower than 0.5 per 100,000. The if statement protects us from the case in which no state satisfies the condition

```{r}
ind <- which.min(murder_rate)

if(murder_rate[ind] < 0.5){
  print(murders$state[ind]) 
} else{
  print("No state has murder rate that low")
}
```

with rate 0.25

```{r}
if(murder_rate[ind] < 0.25){
  print(murders$state[ind]) 
} else{
  print("No state has a murder rate that low.")
}
```

ifelse is also very useful
three arguments; logical, 2 possible answers
TRUE : second argument value is returned, FALSE: third argument value is returned

```{r}
a <- 0
ifelse(a > 0, 1/a, NA)
```

it works on vectors.
examines each entry of the logical vector and returns elements

```{r}
a <- c(0,1,2,-4,5)
result <- ifelse(a > 0, 1/a, NA)
result
```

example of how this function can be used to replace missing values with zeros

```{r}
data(na_example)
no_nas <- ifelse(is.na(na_example), 0, na_example)
sum(is.na(no_nas))
```

two other useful functions : any and all.
any function takes vector of logicals and returns TRUE if any TRUE
function all takes vector of logicals and returns TRUE if all TRUE

```{r}
z<- c(TRUE, TRUE, FALSE)
any(z)
all(z)
```

# 4.2 Defining functions

you'll need to perform same operations over and over
ex) averages
we can compute average of vector x using sum and length functions: sum(x)/length(x)
more efficient to write a function

```{r}
avg <- function(x){
  s <- sum(x)
  n <- length(x)
  s/n
}
```

```{r}
x <- 1:100
identical(mean(x), avg(x))
```

Variables defined inside a function are not saved in the workplace
while we use s and n when we call avg, values are created and changed only during the call

```{r}
s<-3
avg(1:10)
s
```

s is still 3 after we call avg.
functions are objects so we assign them to variable names with <-

```{r}
my_function <- function(VARIABLE_NAME) {
  perform operations on VARIABLE_NAME and calculate VALUE
  VALUE
}
```

the functions you define can have multiple arguments as well as default values

```{r}
avg <- function(x, arithmetic = TRUE){
  n <- length(x)
  ifelse(arithmetic, sum(x)/n, prod(x)^(1/n))
}
```

# 4.3 namespaces

two packages use the same name for two different functions. DPLYR and STATS package define a filter function.

```{r}
library(dplyr)

```

so what does R do when we type filter?
these function live in different namespaces
R will follow certain order when searching for a function

```{r}
search()
```

The first entry in this list is the global environment which includes all the objects you define

stats filter?
you can use ::

```{r}
stats::filter
```

if you want to be sure we use the dplyr filter we use

```{r}
dplyr::filter
```

if we want to use function in a pacakge without loading the entire package, we can use the double colon as well

