---
title: "R Notebook"
output: html_notebook
---

# Chapter 14. Random Variables.

Being able to quantify the uncertainty introduced by randomness is one of the most important jobs of a data analyst.

In this chapter, we introduce random variables and their properties starting with their application to games of chance.

## 14.1 Random Variables

Random variables: numeric outcomes resulting from random process.

```{r}
beads <- rep( c("red", "blue"), times = c(2,3))
X <- ifelse(sample(beads, 1) == "blue", 1, 0)
```

X: random variable.

```{r}
ifelse(sample(beads, 1) == "blue", 1, 0)
ifelse(sample(beads, 1) == "blue", 1, 0)
ifelse(sample(beads, 1) == "blue", 1, 0)
```

Sometimes 0, sometimes 1.

## 14.2 Sampling models

Many data generation procedures can be modeled quite well as draws from an urn.

In experimental research, we often assume that the individual organisms we are studying are a random sample from larger population.

Suppose a very small casino hires you to consult on whether they should set up roulette wheels.
1000 ppl will play, only game you can play is to bet on red or black. Casino wants to know the chance of losing $.

S: random variable, represent the casino's total winnings

```{r}
color <- rep(c("Black", "Red", "Green"), c(18, 18, 2))
```

red comes up; gambler wins, -$1
Ohterwise, the casino wins a dollar and we draw $1

```{r}
n <- 1000
X <- sample(ifelse(color == "Red", -1, 1),  n, replace = TRUE)
X[1:10]
```

We know the proportions of 1s and -1s, we can generate draws with this one line of code:

```{r}
X <- sample(c(-1,1), n, replace = TRUE, prob=c(9/19, 10/19))
```

We call this a 'sampling model since we are modeling the random behavior of roulette with sampling of draws from an urn.

S: sum of 1000 independent draws

```{r}
X <- sample(c(-1,1), n, replace = TRUE, prob=c(9/19, 10/19))
S <- sum(X)
S
```

## 14.3 The probability distribution of a random variable

S changes every time.

Probability distribution of a random variable; tells us probability of observed value falling at any given interval.

F(a)=Pr(S<=a) -> able to answer probability of events

F: random variable's _Distribution Function_

experiment of having 1,000 people play roulette, over and over, specifically B=10000 times:

```{r}
n <- 1000
B <- 10000
roulette_winnings <- function(n){
  X <- sample(c(-1,1), n, replace = TRUE, prob=c(9/19, 10/19))
  sum(X)
}
S <- replicate(B, roulette_winnings(n))
```

```{r}
# mean(S<=a) : how often did we get sums less than or equal to a?

mean(S<0)
```

We can visualize distribution of S by creating a histogram showing F(b)-F(a) for several intervals (a, b]

Distribution appears to be approximately normal.

If the distribution is normal, then all we need to define the distribution is the average and the standard deviation.

this average: 'Expected value', standard deviation: 'Standard eror' of the random variable S.

Statistical theory provides a way to derive the distribution of random variables defined as independent random draws from an urn.

function 'dbinom' and 'pbinom'

To compute Pr(S<0), we note that Pr(S<0) = Pr((S+n)/2<(0+n)/2)

We can use pbinom to compute Pr(S<=0)

```{r}
n <- 1000
pbinom(n/2, size = n, prob = 10/19)
```

Because this is a discrete probability function, to get Pr(S<0) rather than Pr(S<=0),

```{r}
pbinom(n/2-1, size = n, prob = 10/19)
```

## 14.4 Distributions versus probability distributions

F(a): function that tells us what proportion of the list is less than or equal to a.
They are useful summaries when the distribution is approximately normal. We define the average and standard deviation:

```{r}
m <- sum(x)/length(x)
s <- sqrt(sum((x - m)^2) / length(x))
```

A random variable X has a distribution function, and we do not need a list of numbers to define this.

If X is defined by drawing from an urn with numbers in it, then there's a list of numbers inside the urn: distribution of list is probability distribution of X and the average and standard deviation of list are the expected value and standard error of the random variable.

Monte Carlo simulation; outcomes are a list of numbers. The distribution of this list will be a good approximation of the probability distribution of X.

## 14.5 Notation for random variables

upper case letters; denote random variables.
lower case letters; observed values

X<=x
X: random variable, x: arbitrary value and not random. X represent the number on a die role and x represent an actual value we see.

## 14.6 The expected value and standard error

Mathematical theory that lets us approximate the probability distributions for the sum of draws:

Expectd value : E[X]

A random variable will vary around its expected value in a way that if you take the average of many, many draws, the average of the draws will approximate the expected value, getting closer and closer the more draws you take.

->>>> Expected value of a random variable defined by one draw is the average of the numbers in the urn. 

```{r}
B <- 10^6
x <- sample(c(-1,1), B, replace = TRUE, prob=c(9/19, 10/19))
mean(x)
```

If the urn has two possible outcomes; a and b, with proportions p and 1-p

E[X] = ap+b(1-p)

Reason we define E[X] : useful for approximating the probability distributions of sum, which is useful for describing the distribution of averages and proportions.

E[X] of the sum of the draws: number of draws x average of the numbers in the urn.

How different can one observation be from the expected value?

Standard error (SE) gives us an idea of the size of the variation around the expected value: SE[X]

__For independent draws__

square(number of draws) x standard deviation of the numbers in the urn

절대값 (b-a)  x 루트(p(1-p))

In our roulette example,

```{r}
2 * sqrt(90)/19
```

The standard error tells us the typical difference between a random variable and its expectation.

```{r}
n <- 1000
sqrt(n) * 2 * sqrt(90)/19
```

As a result, when 1,000 people bet on red, the casino is expected to win $50 with a standard error of $32. It therefore seems like a safe bet. But we still haven’t answered the question: how likely is it to lose money? Here the CLT will help.

+ Exact probability calculations for the casino winnings can be performed with the binomial distribution

### 14.6.1 Population SD versus the sample SD

The standard deviation of a list x is defined as the square root of the average of the squared differences:

```{r}
library(dslabs)
x <- heights$height
m <- mean(x)
s <- sqrt(mean((x-m)^2))
```

sd function ; different result

```{r}
identical(s, sd(x))
s-sd(x)
```

sd function R does not return the sd of the list, but rather uses a formula that estimates standard deviations of a population from a random sample and divide the sum of the squares by N-1:

```{r}
n <- length(x)
s-sd(x)*sqrt((n-1) / n)
```

actual standard deviation:

```{r}
sqrt(mean((x-m)^2))
```

We sometimes use the sd function when we really want the actual SD: list size is big, these two are practically equivalent

