---
title: "R Notebook"
output: html_notebook
---

## 14.7  Central Limit Theorem
The Central Limit Theorem (CLT) tells us that when the number of draws, also called the sample size, is large, the probability distribution of the sum of the independent draws is approximately normal.

```{r}
n <- 1000
B <- 10000
roulette_winnings <- function(n){
  X <- sample(c(-1,1), n, replace = TRUE, prob=c(9/19, 10/19))
  sum(X)
}
S <- replicate(B, roulette_winnings(n))
```

The Central Limit Theorem (CLT) tells us that the sum \(S\) is approximated by a normal distribution.

E(X) and SE(X)
```{r}
n * (20-18)/38 
sqrt(n) * 2 * sqrt(90)/19 
```

theoretical value match

```{r}
mean(S)
sd(S)
```

```{r}
mu <- n * (20-18)/38
se <-  sqrt(n) * 2 * sqrt(90)/19 
pnorm(0, mu, se)
```

```{r}
mean(S < 0)
```

## 14.7.1  How large is large in the Central Limit Theorem?

Large is a relative term. In many circumstances as few as 30 draws is enough to make the CLT useful.

When the probability of a success is very low: the Poisson distribution is more appropriate.

function 'dpois', 'ppois', generate varaibles using 'rpois'

# 14.8 Statistical properties of averages

1. The expected value of the sum of random variables is the sum of each random variable’s expected value.

2. The expected value of a non-random constant times a random variable is the non-random constant times the expected value of a random variable.

3. The square of the standard error of the sum of independent random variables is the sum of the square of the standard error of each random variable. 

4. The standard error of a non-random constant times a random variable is the non-random constant times the random variable’s standard error.

5. If X is a normally distributed random variable, then if a and b are non-random constants, aX+b is also a normally distributed random variable. All we are doing is changing the units of the random variable by multiplying by a then shifting the center by b.


