---
title: "R Notebook"
output: html_notebook
---
# 5 The Tidyverse
# 5.1 Tidy Data

data table is in tidy format: each row represents one observation and columns represent different variables

```{r}
library(dslabs)
head(murders)
```
tidy!


```{r}
#>        country year fertility
#> 1      Germany 1960      2.41
#> 2  South Korea 1960      6.16
#> 3      Germany 1961      2.44

```

This wasn't this format originally. reshaped for dslabs package

```{r}
#>       country 1960 1961 1962 1963 1964 1965
#> 1     Germany 2.41 2.44 2.47 2.49 2.49 2.48
#> 2 South Korea 6.16 5.99 5.79 5.57 5.36 5.16
```

The same information is provided, but there are two differences
1) each row includes several observations
2) one of the variables, year, is stored in the header
to be used, data should be reshaped into 'tidy' format

# 5.2 Exercises

Q1. Examine the built-in dataset co2. Which of the following is true:

A. co2 is tidy data: it has one year for each row.

B. co2 is not tidy: we need at least one column with a character vector.

C. co2 is not tidy: it is a matrix instead of a data frame.

D. co2 is not tidy: to be tidy we would have to wrangle it to have three columns (year, month and value), then each co2 observation would have a row.

```{r}
co2
# A1. D

```

Q2. Examine the built-in dataset ChickWeight. Which of the following is true:

A. ChickWeight is not tidy: each chick has more than one row.

B. ChickWeight is tidy: each observation (a weight) is represented by one row. The chick from which this measurement came from/ is one the variables.

C. ChickWeight is not a tidy: we are missing the year column.

D. ChickWeight is tidy: it is stored in a data frame.

```{r}
ChickWeight
str(ChickWeight)
# A2. B
```

Q3. Examine the built-in dataset BOD. Which of the following is true:

A. BOD is not tidy: it only has six rows.

B. BOD is not tidy: the first column is just an index.

C. BOD is tidy: each row is an observation with two values (time and demand)

D. BOD is tidy: all small datasets are tidy by definition.

```{r}
BOD
# A3: C
```

Q4. Which of the following built-in datasets is tidy (you can pick more than one):

A. BJsales

B. EuStockMarkets

C. DNase

D. Formaldehyde

E. Orange

F. UCBAdmissions 

```{r}
BJsales
EuStockMarkets
DNase
Formaldehyde
Orange
UCBAdmissions

# A4. C, D, E, F
```

# 5.3 Manipulating Data frames

The Dplyr package from the tidyverse introduces functions with easy names that pwork with data frames ex) mutate: change the data table by adding a new column
filter: filter the data table to a subset of rows
select: to subset the data by selecting specific columns

# 5.3.1 Adding a column with mutate

add murder rates to murders dataframe
'mutate' takes data frame as first argument, name and value of the variable as a second argument using name=values

```{r}
library(dplyr)
library(dslabs)
data("murders")
murders <- mutate(murders, rate = total / population * 100000)
```


we are using 'total' and 'population, which are not defined in workspace. WHY IS THERE NO ERROR?
functions in dplyr package know to look for variables in the data frame in first argument
total = murders$total
new column is added!

```{r}

head(murders)

```

* Though we overwrote the 'murders' object, this does not change the object that loaded with data(murders).

# 5.3.2 Subsetting with filter

Suppose we want to filter data table only to show the entries for which the murder rate is lower than 0.71
'filter' : takes data table as first argument, conditional statement as the second.
we can use the unquoted variable names from murders

```{r}

filter(murders, rate <=0.71)

```

# 5.3.3 Selecting columns with select

If we want to view just a few columsn, we can use the dplyr select function

```{r}
new_table <- select(murders, state, region, rate)
filter(new_table, rate <= 0.71)
```

the first argument is an object, and next are variable names

# 5.4 Exercises

1. Load the dplyr package and the murders dataset

```{r}

library(dplyr)
library(dslabs)

data(murders)
```

You can add columns using the dplyr function mutate. This function is aware of the column names and inside the function you can call them unquoted:

```{r}
murders <- mutate(murders, population_in_millions = population / 10^6)
```

Use the function mutate to add a murders column named rate with the per 100,000 murder rate as in the example code above. Make sure you redefine murders as done in the example code above ( murders <- [your code]) so we can keep using this variable.

```{r}

# A1. 

names(murders)
murders <- mutate(murders,per_tenthousand_rate= total/population * 10^5)
murders

```

Q2. If rank(x) gives you the ranks of x from lowest to highest, rank(-x) gives you the ranks from highest to lowest. Use the function mutate to add a column rank containing the rank, from highest to lowest murder rate. Make sure you redefine murders so we can keep using this variable.

```{r}

murders <- mutate(murders, rank=rank(-per_tenthousand_rate))
murders

```


Q3. With dplyr, we can use select to show only certain columns. For example, with this code we would only show the states and population sizes:

```{r}

select(murders, state, population) %>% head()

```

Use select to show the state names and abbreviations in murders. Do not redefine murders, just show the results.

```{r}

select(murders, state, abb)

```

Q4. The dplyr function filter is used to choose specific rows of the data frame to keep. Unlike select which is for columns, filter is for rows. For example, you can show just the New York row like this:

```{r}
filter(murders, state == "New York")
```

You can use other logical vectors to filter rows.

Use filter to show the top 5 states with the highest murder rates. After we add murder rate and rank, do not change the murders dataset, just show the result. Remember that you can filter based on the rank column.

```{r}

filter(murders, rank<=5 )

```

Q5. We can remove rows using the != operator. For example, to remove Florida, we would do this:

```{r}
no_florida <- filter(murders, state != "Florida")
```

Create a new data frame called no_south that removes states from the South region. How many states are in this category? You can use the function nrow for this.

```{r}

no_south <- filter(murders, region != "South")
nrow(no_south)

```

Q6. We can also use %in% to filter with dplyr. You can therefore see the data from New York and Texas like this:

```{r}

filter(murders, state %in% c("New York", "Texas"))

```

Create a new data frame called murders_nw with only the states from the Northeast and the West. How many states are in this category?

```{r}

murders_nw <- filter(murders, region %in% c("Northeast", "West"))
nrow(murders_nw)

```

Q7. Suppose you want to live in the Northeast or West and want the murder rate to be less than 1. We want to see the data for the states satisfying these options. Note that you can use logical operators with filter. Here is an example in which we filter to keep only small states in the Northeast region.

```{r}
filter(murders, population < 5000000 & region == "Northeast")
```

Make sure murders has been defined with rate and rank and still has all states. Create a table called my_states that contains rows for states satisfying both the conditions: it is in the Northeast or West and the murder rate is less than 1. Use select to show only the state name, the rate and the rank.

```{r}

my_states <- filter(murders, region %in% c("Northeast", "West") &per_tenthousand_rate <1)
my_states
select(my_states, state, per_tenthousand_rate, rank)

```

# 5.5 The pipe: %>%

We can perform a series of operations, by sending the results of one function to another using pipe operator %>%
it looks like
original data -> select -> filter
```{r}
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
```
equivalent to the two lines of code above!
Pipe sends the result of the left side of the pipe to be the first argument on the function on the right side of the pipe

```{r}
16 %>% sqrt()
```

We can continues to pipe values along
```{r}
16 %>% sqrt() %>% log2()
```

the above statement is same with log2(sqrt(16))
```{r}
16 %>% sqrt() %>% log(base = 2)
```

We no longer need to specify the first argument when using the pipe with data frames and dplyr 
```{r}
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
```

murders is the first argument of the select function, and the new data frame (formerly new_table) is the first argument of the filter function.
Pipe works well with functions where the first argument is the input data

# 5.6 Exercises

Q1. The pipe %>% can be used to perform operations sequentially without having to define intermediate objects. Start by redefining murder to include rate and rank.

```{r}
murders <- mutate(murders, rate =  total / population * 100000, rank = rank(-rate))
```

previous solution, we did

```{r}
my_states <- filter(murders, region %in% c("Northeast", "West") & rate < 1)

select(my_states, state, rate, rank)
```

The pipe %>% permits us to perform both operations sequentially without having to define an intermediate variable my_states. We therefore could have mutated and selected in the same line like this

```{r}
mutate(murders, rate =  total / population * 100000, rank = rank(-rate)) %>%
  select(state, rate, rank)
```

Notice that select no longer has a data frame as the first argument. The first argument is assumed to be the result of the operation conducted right before the %>%.

Repeat the previous exercise, but now instead of creating a new object, show the result and only include the state, rate, and rank columns. Use a pipe %>% to do this in just one line.

```{r}

filter(murders, region %in% c("Northeast", "West") & per_tenthousand_rate < 1) %>% select(state, rate, rank)

```

Q2. Reset murders to the original table by using data(murders). Use a pipe to create a new data frame called my_states that considers only states in the Northeast or West which have a murder rate lower than 1, and contains only the state, rate and rank columns. The pipe should also have four components separated by three %>%. The code should look something like this:

```{r}
my_states <- murders %>%
  mutate SOMETHING %>% 
  filter SOMETHING %>% 
  select SOMETHING
```

```{r}
data(murders)
my_states <- murders  %>%
  mutate(murder_rate=total/population*10^5, rank=rank(-murder_rate)) %>%
  filter(region %in% c("Northeast", "West") & murder_rate<1) %>% select(state, murder_rate, rank)
my_states
```
